Supplementary Information: Drought sensitivity is not a species trait: Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments

### 2025-10-10 doc/manuscript_hkk_drought_sensitivity.Rmd
updated files with small cosmetic fixes

**Number of Words in Abstract:** 146  
**Number of Words in Main Text:** 5054   
**Number of References:** 55  
**Number of Figures:** 5  
**Number of Tables:** 1  
**Number of Text boxes:** 0

@@

<!-- The increment timeseries using dendrometer band measurements revealed plot- and species-level patterns in interannual growth variation (Fig `r fig_drought_responses`a).  -->
The three climatically distinct droughts can be characterised as follows: 2010 was a moderate, dry-season drought, 2015 a severe wet-season drought and 2020, a moderate drought during the dry-to-wet transition (Figs `r fig_droughts`, S1, S2).

@@

Species median sensitivities and individual tree sensitivities to the three droughts were not correlated with each other, supporting prediction (*i*) (Fig S5, Fig S6). 
Median annual increment across years varied among the `r nrow(tree.time.sp)` species analysed, with *`r tree.time.sp$spfull[1]`* the lowest (`r tree.time.sp$inc_annual_med[1]`<!-- $\pm$ `r tree.time.sp$inc_annual_sd[1]`--> mm) and *`r tree.time.sp$spfull[nrow(tree.time.sp)]`* the highest (`r tree.time.sp$inc_annual_med[nrow(tree.time.sp)]`<!-- $\pm$ `r tree.time.sp$inc_annual_sd[nrow(tree.time.sp)]`--> mm).
In 2010, species-level sensitivities ranged from `r sens.sp[sens.sp$yr==2010,]$sens_med[1]` for *`r sens.sp[sens.sp$yr==2010,]$spfull[1]`* to `r sens.sp[sens.sp$yr==2010,]$sens_med[nrow(sens.sp)/3]` for *`r sens.sp[sens.sp$yr==2010,]$spfull[nrow(sens.sp)/3]`*. 
In 2015, these ranged from `r sens.sp[sens.sp$yr==2015,]$sens_med[1]` for *`r sens.sp[sens.sp$yr==2015,]$spfull[1]`* to `r sens.sp[sens.sp$yr==2015,]$sens_med[nrow(sens.sp)/3]` for *`r sens.sp[sens.sp$yr==2015,]$spfull[nrow(sens.sp)/3]`*, and in 2020 from `r sens.sp[sens.sp$yr==2020,]$sens_med[1]` for *`r sens.sp[sens.sp$yr==2020,]$spfull[1]`* to `r sens.sp[sens.sp$yr==2020,]$sens_med[nrow(sens.sp)/3]` for *`r sens.sp[sens.sp$yr==2020,]$spfull[nrow(sens.sp)/3]`*. 
Species had more negative sensitivities to the 2015 drought than the 2010 or 2020 droughts, in alignment with median sensitivities across all trees <!--(for 2010 and 2015 Pearson's correlation coefficient r = `r round(sens.sp.cor_10_15[4]$estimate, 2)`, p = `r round(sens.sp.cor_10_15[3]$p.value,2)`, for 2015 and 2020, r = `r round(sens.sp.cor_15_20[4]$estimate, 2)`, p = `r round(sens.sp.cor_15_20[3]$p.value,2)`, and for 2010 and 2020, r = `r round(sens.sp.cor_10_20[4]$estimate, 2)`, p = `r round(sens.sp.cor_10_20[3]$p.value,2)`, -->(Fig S5). Median species sensitivities to the 2010 and 2020 droughts were variable, but no species had higher median increments than a typical year in 2015. <!--KAT reworded here because Mitrephora thorelii was very slightly positive-->

@@

<!----Limitations in filling gap/ comparing only two droughts---->
Our analysis of 30 species across size classes and habitats spans realistic spatial and ecological variation beyond the scope of previous analyses, but we are limited in disentangling the effects of drought characteristics because of limited sample size. 
<!-- Long-timescale analyses with tree rings are limited by the fact that species with rings are few, represent specific strategy to water use and growth - and often affiliated to drier microenvironments. 
Tape measurements of growth from long-term ecological dynamics plots can be representative of species strategies but lack precision to parse signal from noise; error for each tape measure is 1mm, error for growth is at least 2mm while median annual growth across trees in HKK is ~1.4 mm and measuring deviation from this annual growth needs more precision.  -->
<!-- Traditional methods may only be capturing the most dramatic effects - the most severe droughts that affect ring-forming species or produce large-enough declines in fast-growing individuals.  
Further, although estimates of woody growth from dendroband timeseries can have measurement biases because of stem water shrinkage and bark peeling [@chitra-tarak_it_2015], our results are robust to simplifying assumptions that account for these biases (Fig S--).
However, given the relative shorter duration of the timeseries, the effort required to collect these data, and the unpredictability of drought events, our analysis only included two specific droughts.-->
The three droughts analyzed differed in multiple aspects, including seasonal timing, magnitude and duration of meteorological anomalies, and severity as classified with SPEI (Table 1, Fig. `r fig_droughts`). 
<!-- This is a central finding of our study.  -->
<!-- Quantitative analyses parsing out the specific effects of drought characteristics (e.g., timing, degree of anomaly) are currently impossible.  I don't understand this, as we have done this, no? -->
Our finding that a severe wet-season drought in 2015 had greater impact on woody growth than dry-season drought in 2010 and dry-to-wet-season drought in 2020 is somewhat contrary to previous findings. 
Tree ring studies from this site [@vlam_temperature_2014;@anderson-teixeira_joint_2022] and across the tropics [@zuidema_tropical_2022] have asserted that tree growth is most sensitive to dry-season precipitation. 
However, the strength of the ecological response to the 2015 wet-season drought (Fig. `r fig_droughts`) does not align with this conslcusion. 
Generalization of tropical tree growth responses likely requires fine-scale monitoring of growth (i.e., dendrometer band measurements or dendrochronology when possible) on species that span strategies and over different timings and intensities of droughts, together with concurrent measures of soil water avaialability. 
Advancing efforts to collect these data is critical to assessment and prediction of tropical forest responses to changing precipitation patterns.
<!-- [is this necessary?]  Studies in temperate forests, where extensive tree-ring records are available, have confirmed that drought characteristics including seasonal timing, duration, and severity affect tree growth responses [@gao_dynamic_2018; @dorangeville_drought_2018] and that effects of species, species traits, tree size, and microenvironment can differ across droughts [@mcgregor_tree_2021].
As tropical dendrometer band records continues to grow, so will the potential to better understand how specific drought characteristics interact with species traits and microenironment to shape growth respones.  -->

@@

<!----Contributions beyond---->
Despite consistent observations of plot-level growth declines across drought events, we show that this scale obscures a variety of species- and tree-level ecological responses that are heterogeneous across droughts. 
An important implication is that drought responses inferred from tree-ring records derived from the few species known to form reliable, dateable annual rings (n=4 at HKK; @vlam_temperature_2014) are unlikely to be representative of the drought responses of the entire community.
For example, one of the few annual-ring-forming species at HKK, deciduous *Afzelia xylocarpa*, had the most distinct response across the three events; it's growth nearly doubled in the 2010 and 2020 droughts but was severely reduced in the wet-season drought (Fig. S5, Table S1).
Extrapolating from such dendrochronological data, therefore, could lead to important errors if attributed to other species, potentially misrepresenting drought impacts conservation or silviculture, e.g. *Dipterocarpus alatus*, a conservation and commercial priority species, presents a different drought response (Table S1).
More generally, tree-ring records likely present a biased picture of drought responses, as ring formation is caused by dry-season dormancy (often associated with deciduous leaf habit) and sampling often targets large trees with exposed crowns and dry microenvironments [@speer_fundamentals_2010]-- all factors shown here to <!--directly and interactively -->influence drought sensitivity.
<!-- More generally, past analyses of tree ring and census estimates of drought sensitivity in seasonally dry tropical forests may have been limited towards fast growing (measurable using tape) or few dry-associated species (that form annual tree rings) - and therefore capturing only the more dramatic effects. -->
Our analyses reveal a complex heterogeneity of drought responses, including the contrasting influence of deciduousness in three different droughts, showing that ecological and evolutionary strategies can make particular species and trees resistant to some droughts but vulnerable to others. 
These distinct responses of species and individuals to droughts suggests potential biological insurance effects [@loreau_biodiversity_2021] among tropical tree responses, in alignment with recent global analyses [@liu_drought_2022; @zuidema_pantropical_2025; @langan_amazon_2025].

>>>>

**Number of Words in Abstract:** 146  
**Number of Words in Main Text:** 5054   
**Number of References:** 58  
**Number of Figures:** 5  
**Number of Tables:** 1  
**Number of Text boxes:** 0

@@

<!-- The increment timeseries using dendrometer band measurements revealed plot- and species-level patterns in interannual growth variation (Fig `r fig_drought_responses`a).  -->
The three climatically distinct droughts can be characterised as follows: 2010 was a moderate, dry-season drought, 2015 a severe wet-season drought and 2020, a moderate drought during the dry-to-wet transition (Figs `r fig_droughts`, S2, S3).

@@

Species median sensitivities and individual tree sensitivities to the three droughts were not correlated with each other, supporting prediction (*i*) (Fig S6, Fig S7). 
Median annual increment across years varied among the `r nrow(tree.time.sp)` species analysed, with *`r tree.time.sp$spfull[1]`* the lowest (`r tree.time.sp$inc_annual_med[1]`<!-- $\pm$ `r tree.time.sp$inc_annual_sd[1]`--> mm) and *`r tree.time.sp$spfull[nrow(tree.time.sp)]`* the highest (`r tree.time.sp$inc_annual_med[nrow(tree.time.sp)]`<!-- $\pm$ `r tree.time.sp$inc_annual_sd[nrow(tree.time.sp)]`--> mm).
In 2010, species-level sensitivities ranged from `r sens.sp[sens.sp$yr==2010,]$sens_med[1]` for *`r sens.sp[sens.sp$yr==2010,]$spfull[1]`* to `r sens.sp[sens.sp$yr==2010,]$sens_med[nrow(sens.sp)/3]` for *`r sens.sp[sens.sp$yr==2010,]$spfull[nrow(sens.sp)/3]`*. 
In 2015, these ranged from `r sens.sp[sens.sp$yr==2015,]$sens_med[1]` for *`r sens.sp[sens.sp$yr==2015,]$spfull[1]`* to `r sens.sp[sens.sp$yr==2015,]$sens_med[nrow(sens.sp)/3]` for *`r sens.sp[sens.sp$yr==2015,]$spfull[nrow(sens.sp)/3]`*, and in 2020 from `r sens.sp[sens.sp$yr==2020,]$sens_med[1]` for *`r sens.sp[sens.sp$yr==2020,]$spfull[1]`* to `r sens.sp[sens.sp$yr==2020,]$sens_med[nrow(sens.sp)/3]` for *`r sens.sp[sens.sp$yr==2020,]$spfull[nrow(sens.sp)/3]`*. 
Species had more negative sensitivities to the 2015 drought than the 2010 or 2020 droughts, in alignment with median sensitivities across all trees <!--(for 2010 and 2015 Pearson's correlation coefficient r = `r round(sens.sp.cor_10_15[4]$estimate, 2)`, p = `r round(sens.sp.cor_10_15[3]$p.value,2)`, for 2015 and 2020, r = `r round(sens.sp.cor_15_20[4]$estimate, 2)`, p = `r round(sens.sp.cor_15_20[3]$p.value,2)`, and for 2010 and 2020, r = `r round(sens.sp.cor_10_20[4]$estimate, 2)`, p = `r round(sens.sp.cor_10_20[3]$p.value,2)`, -->(Fig S6). Median species sensitivities to the 2010 and 2020 droughts were variable, but no species had higher median increments than a typical year in 2015. <!--KAT reworded here because Mitrephora thorelii was very slightly positive-->

@@

<!----Limitations in filling gap/ comparing only two droughts---->
Our analysis of 30 species across size classes and habitats spans realistic spatial and ecological variation beyond the scope of previous analyses, but we are limited in disentangling the effects of drought characteristics because of limited sample size. 
<!-- Long-timescale analyses with tree rings are limited by the fact that species with rings are few, represent specific strategy to water use and growth - and often affiliated to drier microenvironments. 
Tape measurements of growth from long-term ecological dynamics plots can be representative of species strategies but lack precision to parse signal from noise; error for each tape measure is 1mm, error for growth is at least 2mm while median annual growth across trees in HKK is ~1.4 mm and measuring deviation from this annual growth needs more precision.  -->
<!-- Traditional methods may only be capturing the most dramatic effects - the most severe droughts that affect ring-forming species or produce large-enough declines in fast-growing individuals.  
Further, although estimates of woody growth from dendroband timeseries can have measurement biases because of stem water shrinkage and bark peeling [@chitra-tarak_it_2015], our results are robust to simplifying assumptions that account for these biases (Fig S--).
However, given the relative shorter duration of the timeseries, the effort required to collect these data, and the unpredictability of drought events, our analysis only included two specific droughts.-->
The three droughts analyzed differed in multiple aspects, including seasonal timing, magnitude and duration of meteorological anomalies, and severity as classified with SPEI (Table 1, Fig. `r fig_droughts`). 
<!-- This is a central finding of our study.  -->
<!-- Quantitative analyses parsing out the specific effects of drought characteristics (e.g., timing, degree of anomaly) are currently impossible.  I don't understand this, as we have done this, no? -->
Our finding that a severe wet-season drought in 2015 had greater impact on woody growth than dry-season drought in 2010 and dry-to-wet-season drought in 2020 is somewhat contrary to previous findings. 
Tree ring studies from this site [@vlam_temperature_2014;@anderson-teixeira_joint_2022] and across the tropics [@zuidema_tropical_2022] have asserted that tree growth is most sensitive to dry-season precipitation. 
However, the strength of the ecological response to the 2015 wet-season drought (Fig. `r fig_droughts`) does not align with this conclusion. 
Generalization of tropical tree growth responses likely requires fine-scale monitoring of growth (i.e., dendrometer band measurements or dendrochronology when possible) on species that span strategies and over different timings and intensities of droughts, together with concurrent measures of soil water availability. 
Advancing efforts to collect these data is critical to assessment and prediction of tropical forest responses to changing precipitation patterns.
<!-- [is this necessary?]  Studies in temperate forests, where extensive tree-ring records are available, have confirmed that drought characteristics including seasonal timing, duration, and severity affect tree growth responses [@gao_dynamic_2018; @dorangeville_drought_2018] and that effects of species, species traits, tree size, and microenvironment can differ across droughts [@mcgregor_tree_2021].
As tropical dendrometer band records continues to grow, so will the potential to better understand how specific drought characteristics interact with species traits and microenironment to shape growth respones.  -->

@@

<!----Contributions beyond---->
Despite consistent observations of plot-level growth declines across drought events, we show that this scale obscures a variety of species- and tree-level ecological responses that are heterogeneous across droughts. 
An important implication is that drought responses inferred from tree-ring records derived from the few species known to form reliable, dateable annual rings (n=4 at HKK; @vlam_temperature_2014) are unlikely to be representative of the drought responses of the entire community.
For example, one of the few annual-ring-forming species at HKK, deciduous *Afzelia xylocarpa*, had the most distinct response across the three events; it's growth nearly doubled in the 2010 and 2020 droughts but was severely reduced in the wet-season drought (Fig S6, Table S1).
Extrapolating from such dendrochronological data, therefore, could lead to important errors if attributed to other species, potentially misrepresenting drought impacts conservation or silviculture, e.g. *Dipterocarpus alatus*, a conservation and commercial priority species, presents a different drought response (Table S1).
More generally, tree-ring records likely present a biased picture of drought responses, as ring formation is caused by dry-season dormancy (often associated with deciduous leaf habit) and sampling often targets large trees with exposed crowns and dry microenvironments [@speer_fundamentals_2010]-- all factors shown here to <!--directly and interactively -->influence drought sensitivity.
<!-- More generally, past analyses of tree ring and census estimates of drought sensitivity in seasonally dry tropical forests may have been limited towards fast growing (measurable using tape) or few dry-associated species (that form annual tree rings) - and therefore capturing only the more dramatic effects. -->
Our analyses reveal a complex heterogeneity of drought responses, including the contrasting influence of deciduousness in three different droughts, showing that ecological and evolutionary strategies can make particular species and trees resistant to some droughts but vulnerable to others. 
These distinct responses of species and individuals to droughts suggests potential biological insurance effects [@loreau_biodiversity_2021] among tropical tree responses, in alignment with recent global analyses [@liu_drought_2022; @zuidema_pantropical_2025; @langan_amazon_2025].

### 2025-10-10 doc/supplementary_information.Rmd
updated files with small cosmetic fixes

@@

![**Figure S1:** Raw climate data timeseries from 2009 to 2023 from ERA5Land and CHIRPS. Measurements represent daily rolling means over a 30-day time window for precipitation, max temperature and VPD and rolling sum for number of dry days. The shaded areas indicate the three drought years analysed - 2010, 2015 and 2020.](display/climvars_fullseries_remote.png)

![**Figure S2:** Climate anomalies for all variables from 2009 to 2020. Values represent number of standard deviations from the long-term mean for daily rolling mean values. The shaded areas indicate the three drought years analysed - 2010, 2015 and 2020.](display/climvars_anomalies_fullseries_remote.png)

@@

![**Figure S3:** Growth anomalies across the timeseries. Growth anomalies for each year calculated as the number of standard deviations in growth each year from mean growth across all years and summarised for plot and species.](display/growth_anom_fullseries.png)

@@

![**Figure S4:** Species sensitivities in drought years 2010, 2015 and 2020. Boxplots represent 25th, 50th (median) and 75th percentile of sensitivity for each species in each year and whiskers represent 95% CIs. All individual sensitivities represented by jittered points.](display/Fig_SI_species_sensitivities.png)

![**Figure S5:** Correlations of species sensitivities across the two years. Panel a shows correlation of species median sensitivities in the two years and Pearson's correlation coefficient. Colour gradient represents deciduousness scale and dotted line is the 1:1 line. Panel b shows difference in magnitude of species median sensitivites among the two years](display/Fig_SI_sensitivity_cor_dec.png)

@@

![**Figure S6:** Correlation of individual sensitivities across the two years along with Pearson's correlation coefficient and the 1:1 line.](display/sens_corr.png)

@@

![**Figure S7:** Map of trees with dendrobands in HKK, included in analyses in the 2010, 2015 and 2020 datasets](display/dendro_tree_map.png)

![**Figure S8:** Calculated Topographic Wetness Index (TWI) values for the HKK plot](display/twi.png)

![**Figure S9:** Distribution of DBH with Crown Illumination Index (CII) in 2014.](display/dbh_cii.png)

@@

![**Figure S10:** Distribution of raw sensitivity values across years and their relationship with DBH, CII and TWI. Lines are linear model fits.](display/sens_var_rels.png)

![**Figure S11:** Distribution of residuals of species intercept models of sensitivity and their relationship with scaled DBH, TWI and CII and deciduousness. Species intercept model results are reported in main text.](display/residuals.png)

@@

To analyse the influence of microenvironmental variables on growth sensitivity, we first created a Directed Acyclic Graph describing the relationships. To be able to test the DAG with the dataset, we used *dagitty* to derive the conditional dependencies that need to be tested, namely CII $\perp\!\!\!\perp$ TWI and DBH $\perp\!\!\!\perp$ TWI.

![**Figure S12:** Testing correlations of variables across all individuals](display/cond_dep_alltrees.png)

@@

![**Figure S13:** Testing correlations for DBH $\perp\!\!\!\perp$ TWI by species](display/cond_dep_dbh_twi.png)

![**Figure S14:** Testing correlations for CII $\perp\!\!\!\perp$ TWI by species](display/cond_dep_cii_twi.png)

@@

![**Figure S15:** Correlation of modelled species sensitivities with Topographic Position Index](display/FigSI_TPI_preds.png)

@@

![**Figure S16:** Model results from Bayesian causal model of sensitivity as a function of DBH, CII and TPI.](display/Fig4_tpi_3yrs.png)

@@

![**Figure S17:** Model results from Bayesian causal model of sensitivity as a function of DBH, CII and TWI with species random effect only on intercept.](display/FigSI_coefs_nospre.png)

@@

![**Figure S18:** Model results from three independent causal inquiry models of sensitivity following backdoor criteria for total causal effect of DBH, CII and TWI with species random effect on intercept and slope.](display/causal_inquiry.png)

>>>>

## Study site

![**Figure S1:** Representative photographs of the 50-ha long-term monitoring site in Huai Kha Khaeng Thailand, with dendrometer banded trees. Photo credits: left image: Sarayudh Bunyavejchewin, right image: Krishna Anujan](display/site_photos.png)

@@

![**Figure S2:** Raw climate data timeseries from 2009 to 2023 from ERA5Land and CHIRPS. Measurements represent daily rolling means over a 30-day time window for precipitation, max temperature and VPD and rolling sum for number of dry days. The shaded areas indicate the three drought years analysed - 2010, 2015 and 2020. Dry-season in the region is typically from November to April.](display/climvars_fullseries_remote.png)

![**Figure S3:** Climate anomalies for all variables from 2009 to 2020. Values represent number of standard deviations from the long-term mean for daily rolling mean values. The shaded areas indicate the three drought years analysed - 2010, 2015 and 2020.Dry-season in the region is typically from November to April.](display/climvars_anomalies_fullseries_remote.png)

@@

![**Figure S4:** Growth anomalies across the timeseries. Growth anomalies for each year calculated as the number of standard deviations in growth each year from mean growth across all years and summarised for plot and species.](display/growth_anom_fullseries.png)

@@

![**Figure S5:** Species sensitivities in drought years 2010, 2015 and 2020. Boxplots represent 25th, 50th (median) and 75th percentile of sensitivity for each species in each year and whiskers represent 95% CIs. All individual sensitivities represented by jittered points.](display/Fig_SI_species_sensitivities.png)

![**Figure S6:** Correlations of species sensitivities across the two years. The top-left panel shows correlation of species median sensitivities and Pearson's correlation coefficient in the 2010 and 2015, top right shows 2010 and 2020, and bottom left shows 2015 and 2020. Colour gradient represents deciduousness scale and dotted line is the 1:1 line. The bottom right panel shows how species drought sensitivites varies among the three years; the color scheme for droughts matches that in main text figure 1.](display/Fig_SI_sensitivity_cor_dec.png)

@@

![**Figure S7:** Correlation of individual sensitivities across the two years along with Pearson's correlation coefficient and the 1:1 line.](display/sens_corr.png)

@@

![**Figure S8:** Map of trees with dendrobands in HKK, included in analyses in the 2010, 2015 and 2020 datasets](display/dendro_tree_map.png)

![**Figure S9:** Calculated Topographic Wetness Index (TWI) values for the HKK plot](display/twi.png)

![**Figure S10:** Distribution of DBH with Crown Illumination Index (CII) in 2014.](display/dbh_cii.png)

@@

![**Figure S11:** Distribution of raw sensitivity values across years and their relationship with DBH, CII and TWI. Blue lines are linear model fits for statistically significant relationships at p<0.05. Red lines are Friedman's SuperSmoother fits for all.](display/sens_var_rels.png)

![**Figure S12:** Distribution of residuals of species intercept models of sensitivity and their relationship with scaled DBH, TWI and CII and deciduousness. Grey lines are linear model fits for statistically significant relationships at p<0.05. Species intercept model results are reported in main text.](display/residuals.png)

@@

To analyse the influence of microenvironmental variables on growth sensitivity, we first created a Directed Acyclic Graph describing the relationships. 
The consistency of the DAG with the dataset, and our ability build causally interpretable models, is conditional on the numerical independence among some specific variable groups, as implied by the DAG structure.
We identified these "testable implications" or "conditional independences" for this DAG using *dagitty* as CII $\perp\!\!\!\perp$ TWI and DBH $\perp\!\!\!\perp$ TWI.
We tested these correlations using the function *cor()* in R, with the simple criteria that a Pearson's correlation coefficient <0.3 implies independence among variable pairs and causal relevance of the DAG.

![**Figure S13:** Testing correlations of variables across all individuals](display/cond_dep_alltrees.png)

@@

![**Figure S14:** Testing correlations for DBH $\perp\!\!\!\perp$ TWI by species](display/cond_dep_dbh_twi.png)

![**Figure S15:** Testing correlations for CII $\perp\!\!\!\perp$ TWI by species](display/cond_dep_cii_twi.png)

@@

![**Figure S16:** Correlation of modelled species sensitivities with Topographic Position Index](display/FigSI_TPI_preds.png)

@@

![**Figure S17:** Model results from Bayesian causal model of sensitivity as a function of DBH, CII and TPI.](display/Fig4_tpi_3yrs.png)

@@

![**Figure S18:** Model results from Bayesian causal model of sensitivity as a function of DBH, CII and TWI with species random effect only on intercept.](display/FigSI_coefs_nospre.png)

@@

![**Figure S19:** Model results from three independent causal inquiry models of sensitivity following backdoor criteria for total causal effect of DBH, CII and TWI with species random effect on intercept and slope.](display/causal_inquiry.png)
